---
title: "Candida_sc_clust5"
author: "Nathan Morris"
date: "2023-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```


# Cluster 5 Analysis

### Data Import
```{r echo = TRUE, include= TRUE, error = FALSE, warning = FALSE, message=FALSE}
# Data Import
library(Seurat)
library(scater)
library(dplyr)
library(SingleR)
library(celldex)
library(scRNAseq)
library(scran)
library(gridExtra)
library(pheatmap)
library(ess)
library(scCustomize)
library(reticulate)
library(ReductionWrappers)

data_dir <- "/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Candida_Migratory_TCR_Project/Data"
setwd(data_dir)
```

```{r}
seurat_object <- readRDS(paste0(data_dir,"/dataset_post_sc_Analysis.Rds"))
```

## For Reticulate (Python-R wrapper)
For reticulate - python virtual environment activation
This makes a conda enviornment r-reticulate, and then downloades the opentsne and phate packages from python (https://github.com/pavlin-policar/openTSNE/ and https://github.com/KrishnaswamyLab/PHATE).
Therefore, we can use these packages, when we call the r-reticulate package and the ReductionWrappers, which are R language wrappers for these packages.
```{r}
conda_list()
use_condaenv("r-reticulate")    #activate an environment e.g py2_env
py_install("opentsne")
reticulate::py_install("phate", pip=TRUE)
devtools::install_github("KrishnaswamyLab/phateR")
py_install("fa2")

#reticulate::py_discover_config("phate")

```

# Cluster 5 - Identifying Cells when Subsetting

```{r}
Idents(seurat_object) <- "RNA_clusters"
Cluster_Highlight_Plot(seurat_object = seurat_object, cluster_name = c("5"), highlight_color = c("navy"), background_color = "lightgray")
```

# Finding Marker Genes for All Clust 5

```{r}
# Find all markers of cluster 5 against all
Idents(seurat_object) <- "RNA_clusters"
cluster5.markers <- FindMarkers(seurat_object, ident.1 = 5, only.pos = TRUE, min.pct = 0.05, logfc.threshold = 0.05)
#head(cluster5.markers, n = 5)
```

```{r}
#markers <- seurat_object.RNAmarkers[seurat_object.RNAmarkers$cluster == 5,]
DefaultAssay(object = seurat_object) <- "RNA"
markers <- cluster5.markers[order(cluster5.markers$p_val_adj), ]
markers <- markers[(markers$p_val_adj > 0),]

marker_genes <- rownames(markers[1:8,])
marker_genes <- marker_genes[!is.na(marker_genes)]

more_marker_genes <- rownames(markers[1:50,])
more_marker_genes <- more_marker_genes[!is.na(more_marker_genes)]

FeaturePlot(seurat_object, features = marker_genes, ncol = 4, reduction = "umap_rna")

FeaturePlot(seurat_object, features = marker_genes[1:2], ncol = 2, reduction = "umap_rna", split.by = "Infection")
  
DoHeatmap(seurat_object, features = more_marker_genes, raster = FALSE) + NoLegend()
  
#ggsave(paste("Cluster-5", "_Top50_Markers_Heatmap.pdf", sep = ""), width=16, height=16)

DotPlot(seurat_object, features = more_marker_genes) + RotatedAxis() + coord_flip()
```


# Cluster 5 Organ and Infection Distribution
```{r}
table(seurat_object$Organ[which(seurat_object$RNA_clusters == 5)])
table(seurat_object$Infection[which(seurat_object$RNA_clusters == 5)])
```

Cluster 5 mainly consists of Colon and Lung samples. It also shows a significant difference between No and Candida infected samples.


# Cluster 5 - Re-clustering

```{r}
seurat_clust5 <- subset(x = seurat_object, subset = (RNA_clusters == 5))

seurat_clust5 <- seurat_clust5 %>% NormalizeData(verbose = FALSE) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 4000) %>% ScaleData(verbose = TRUE)
so.clust5 <- seurat_clust5 %>% RunPCA(pc.genes = seurat_clust5@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:25) %>% RunTSNE(dims = 1:25) %>% FindNeighbors(dims = 1:25,compute.SNN=TRUE) %>% FindClusters(resolution = c(0.1,0.2,0.4,0.8,1)) %>% DoopenTSNE(reduction_save = "openTSNE",dims_use = 1:25) 

```

```{r}
# Trying to run PHATE and ForceAtlas
#b_seurat_phate <- DoPHATE(b_seurat, reduction_use = "pca", reduction_save = "phate", dims_use = 1:25) 
so.clust5 <- DoForceAtlas2(so.clust5, reduction_use = "pca", reduction_save = "fa2", dims_use = 1:25)
```

## Visualizing Re-clustering

```{r}
# Infection Plots
DimPlot(so.clust5, reduction = "umap", group.by = "Infection")
DimPlot(so.clust5, reduction = "tsne", group.by = "Infection")
DimPlot(so.clust5, reduction = "openTSNE", group.by = "Infection")
DimPlot(so.clust5, reduction = "fa2", group.by = "Infection")

# Resolution 0.4
DimPlot(so.clust5, reduction = "umap", group.by = "RNA_snn_res.0.4", label = TRUE)
DimPlot(so.clust5, reduction = "tsne", group.by = "RNA_snn_res.0.4", label = TRUE)
DimPlot(so.clust5, reduction = "openTSNE", group.by = "RNA_snn_res.0.4", label = TRUE)
DimPlot(so.clust5, reduction = "fa2", group.by = "RNA_snn_res.0.4", label = TRUE)
DimPlot(so.clust5, reduction = "openTSNE", group.by = "RNA_snn_res.0.4", label = TRUE, split.by = "Infection")

# Resolution 0.8
DimPlot(so.clust5, reduction = "umap", group.by = "RNA_snn_res.0.8", label = TRUE)
DimPlot(so.clust5, reduction = "tsne", group.by = "RNA_snn_res.0.8", label = TRUE)
DimPlot(so.clust5, reduction = "openTSNE", group.by = "RNA_snn_res.0.8", label = TRUE)
DimPlot(so.clust5, reduction = "fa2", group.by = "RNA_snn_res.0.8", label = TRUE)
DimPlot(so.clust5, reduction = "openTSNE", group.by = "RNA_snn_res.0.8", label = TRUE, split.by = "Infection")

# Resolution 1.0
DimPlot(so.clust5, reduction = "umap", group.by = "RNA_snn_res.1", label = TRUE)
DimPlot(so.clust5, reduction = "tsne", group.by = "RNA_snn_res.1", label = TRUE)
DimPlot(so.clust5, reduction = "openTSNE", group.by = "RNA_snn_res.1", label = TRUE)
DimPlot(so.clust5, reduction = "fa2", group.by = "RNA_snn_res.1", label = TRUE)
DimPlot(so.clust5, reduction = "openTSNE", group.by = "RNA_snn_res.1", label = TRUE, split.by = "Infection")

# Density Plots
#Plot_Density_Custom(seurat_object = tnk_seurat_v2, features = "Fos")

```

```{r}
so.clust5$subclust_5 <- so.clust5$RNA_snn_res.0.4
so.clust5$subclust_5.08 <- so.clust5$RNA_snn_res.0.8
so.clust5$subclust_5.1 <- so.clust5$RNA_snn_res.1
```

Add Metadata

```{r}
so.clust5[['Mouse_New']] <- so.clust5[['Mouse']]

Idents(so.clust5) <- 'Mouse_New'

so.clust5 <- RenameIdents(so.clust5, `mouse0082` = 'No Infection Male', `mouse0081` = 'Infected Female', `mouse0080` = 'Infected Male')

so.clust5[['Mouse_New']] <- Idents(so.clust5)

so.clust5$Mouse_New <- factor(so.clust5$Mouse_New, levels = c('No Infection Male', 'Infected Female', 'Infected Male'))

```

```{r}
table(so.clust5$subclust_5, so.clust5$Mouse_New)
table(so.clust5$subclust_5.08, so.clust5$Mouse_New)
table(so.clust5$subclust_5.1, so.clust5$Mouse_New)
```

# Adding Alex's Reclustering 12/18

```{r}
alex_dir <- "/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Data Clubs:Meetings/09_01_Alex_Data/20230901_update/Cluster5_Subclustering/"

poo <- readRDS(file = paste0(alex_dir, "Cluster5_processed_seurat.rds"))

#poo1 <- readRDS(file = paste0(alex_dir, "Cluster5_pseudotime_seurat.rds"))

so.clust5$alex_clusters <- poo[match(so.clust5@meta.data[["cell_barcode"]], poo@meta.data[["cell_barcode"]]),]$seurat_clusters

#table(so.clust5$alex_clusters, so.clust5$Clust5_subclust)

```

# Seeing Alex's Clusters on the whole UMAP
```{r}
# Extract common barcodes
common_barcodes <- intersect(rownames(seurat_object_high_clust), rownames(so.clust5))

# Match and assign alex_clusters
seurat_object_high_clust$alex_clust <- so.clust5[match(common_barcodes, so.clust5$cell_barcode), ]$alex_clusters

Idents(object = seurat_object_high_clust) <- "alex_clust"
DimPlot(seurat_object_high_clust, reduction = "umap_rna", label = TRUE, repel = TRUE)
```

## Mouse Proportion in each Cluster

```{r}
#data <- as.data.frame(table(so.clust5$subclust_5.08, so.clust5$Mouse_New))
data <- as.data.frame(table(so.clust5$alex_clusters, so.clust5$Mouse_New))
colnames(data) <- c("Cluster", "Mouse", "Freq")

result <- data %>%
  group_by(Mouse) %>%
  mutate(Percentage = Freq / sum(Freq) * 100) %>%
  ungroup()

ggplot(result, aes(x = factor(Cluster), y = Percentage, fill = Mouse)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "Percentage Subcluster cells from each Mouse",
       x = "Cluster 5 subcluster",
       y = "Percentage of Subcluster cells from Each Mouse") +
  geom_text(aes(label = paste0(round(Percentage), "%")),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 3) +
  theme_minimal()
```


## Organ Proportion in each Cluster

```{r}
#data <- as.data.frame(table(so.clust5$subclust_5.08, so.clust5$Organ))
data <- as.data.frame(table(so.clust5$alex_clusters, so.clust5$Organ))
colnames(data) <- c("Cluster", "Organ", "Freq")

result <- data %>%
  group_by(Organ) %>%
  mutate(Percentage = Freq / sum(Freq) * 100) %>%
  ungroup()

ggplot(result, aes(x = factor(Cluster), y = Percentage, fill = Organ)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "Percentage of Each Organ in Each Cluster",
       x = "Cluster 5 Subcluster",
       y = "Percentage of Subcluster cells in each Organ") +
  geom_text(aes(label = paste0(round(Percentage), "%")),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 3)  +
  theme_minimal()
```


## 
```{r}
DimPlot(so.clust5, reduction = "fa2", group.by = "alex_clusters", label = TRUE)
DimPlot(so.clust5, reduction = "openTSNE", group.by = "alex_clusters", label = TRUE, split.by = "Infection")

DimPlot(so.clust5, reduction = "openTSNE", group.by = "Organ", label = TRUE, split.by = "Infection")
DimPlot(so.clust5, reduction = "openTSNE", group.by = "Mouse_New", label = TRUE, split.by = "Infection")
```

### Subclusters - All Marker Genes

## 12/18

```{r}
Idents(object = so.clust5) <- "subclust_5.08"
so.clust5_allmark_RNA <- FindAllMarkers(so.clust5, assay = "RNA", logfc.threshold = 0.025, min.pct = 0.025, only.pos = TRUE)

so.clust5_allmark_RNA <- so.clust5_allmark_RNA[so.clust5_allmark_RNA$p_val_adj < 0.10,]

so.clust5_allmark_ADT <- FindAllMarkers(so.clust5, assay = "ADT", logfc.threshold = 0.025, min.pct = 0.025, only.pos = TRUE)
```

```{r}
fig_dir <- ("/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Data_share_12_11/")
write.csv(so.clust5_allmark_RNA, paste0(fig_dir, "Cluster_5_Subcluster_DE_Genes_RNA.csv"), row.names=FALSE)
write.csv(so.clust5_allmark_ADT, paste0(fig_dir, "Cluster_5_Subcluster_DE_Protein_ADT.csv"), row.names=FALSE)
```

```{r}
top_10_df_clust5 <- Extract_Top_Markers(marker_dataframe = so.clust5_allmark_RNA, num_genes = 10, data_frame = TRUE,
    rank_by = "avg_log2FC", named_vector = FALSE, make_unique = TRUE)
head(top_10_df_clust5, 10)
#any(duplicated(x = top_5_df$gene))

```

```{r}
DefaultAssay(so.clust5) <- "RNA"
Clustered_DotPlot(so.clust5, features = unique(top_10_df_clust5$gene))

DoHeatmap(so.clust5, features = unique(top_10_df_clust5$gene), raster = FALSE)
```

# Using Alex Clusters

```{r}
Idents(object = so.clust5) <- "alex_clusters"
so.clust5_allmark_RNA_alex <- FindAllMarkers(so.clust5, assay = "RNA", logfc.threshold = 0.025, min.pct = 0.025, only.pos = TRUE)

so.clust5_allmark_RNA_alex <- so.clust5_allmark_RNA_alex[so.clust5_allmark_RNA_alex$p_val_adj < 0.10,]

so.clust5_allmark_ADT_alex <- FindAllMarkers(so.clust5, assay = "ADT", logfc.threshold = 0.025, min.pct = 0.025, only.pos = TRUE)
```

```
fig_dir <- ("/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Data_share_12_11/")
write.csv(so.clust5_allmark_RNA_alex, paste0(fig_dir, "Cluster_5_alex-Subcluster_DE_Genes_RNA.csv"), row.names=FALSE)
write.csv(so.clust5_allmark_ADT_alex, paste0(fig_dir, "Cluster_5_alex-Subcluster_DE_Protein_ADT.csv"), row.names=FALSE)
```


```{r}
top_10_df_clust5_alex <- Extract_Top_Markers(marker_dataframe = so.clust5_allmark_RNA_alex, num_genes = 10, data_frame = TRUE,
    rank_by = "avg_log2FC", named_vector = FALSE, make_unique = TRUE)
head(top_10_df_clust5_alex, 10)
#any(duplicated(x = top_5_df$gene))

```

```{r}
DefaultAssay(so.clust5) <- "RNA"
Clustered_DotPlot(so.clust5, features = unique(top_10_df_clust5_alex$gene))

DoHeatmap(so.clust5, features = unique(top_10_df_clust5_alex$gene), raster = FALSE)
```

# Alex Cluster Marker Literature Search

Cluster 0:

Cluster 1:

Cluster 2:
Mostly Colon
Somewhat higher concentration of the Infected Population
GPR183 - https://bpspubs.onlinelibrary.wiley.com/doi/epdf/10.1111/bph.15311
  - G-protein coupled receptor 2 - 
  - Migration of GPR183 expressing cells to higher cholesterol oxidation derived Oxysterol concentrations

```{r}
AverageExpression(so.clust5, features = "Mki67", group.by = "alex_clusters", assay = "RNA")
```

```{r}

```

# Merging TCR


```{r}
#Merge TCR

so.clust5$clonotype <- IGT27_TCR[match(so.clust5@meta.data[["cell_barcode"]], IGT27_TCR$Barcode),]$clonotype

table(rev(sort(table(so.clust5$clonotype))))

Idents(object = so.clust5) <- "clonotype"

so.clust5$clonotype_frequency <- IGT27_TCR[match(so.clust5@meta.data[["cell_barcode"]], IGT27_TCR$Barcode),]$Clonotype_frequency

table(rev(sort(table(so.clust5$clonotype))))

so.clust5@meta.data[["paired_across_organ"]] <- so.clust5@meta.data[["cell_barcode"]] %in% IGT27_TCR$Barcode[IGT27_TCR$paired_across_organ == TRUE]

Idents(object = so.clust5) <- "paired_across_organ"

so.clust5@meta.data[["paired_across_organ"]]<- as.factor(so.clust5$paired_across_organ)

so.clust5@meta.data[["duplicate"]] <- so.clust5@meta.data[["cell_barcode"]] %in% IGT27_TCR$Barcode[IGT27_TCR$duplicate == TRUE]

Idents(object = so.clust5) <- "duplicate"

so.clust5@meta.data[["duplicate"]]<- as.factor(so.clust5$duplicate)
```

```{r}
DimPlot(so.clust5, reduction = "umap_rna", group.by = "paired_across_organ", label = TRUE, split.by = "subclust_5.08")

DimPlot(so.clust5, reduction = "umap_rna", group.by = "clonotype_frequency", label = TRUE, split.by = "subclust_5.08")
```

```{r}
data <- as.data.frame(table(so.clust5$subclust_5.08, so.clust5$paired_across_organ))
colnames(data) <- c("Subcluster", "Migratory", "Freq")

result <- data %>%
  group_by(Subcluster) %>%
  mutate(Percentage = Freq / sum(Freq) * 100) %>%
  ungroup()

ggplot(result, aes(x = factor(Subcluster), y = Percentage, fill = Migratory)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "Percentage of Migratory Cells in Each Subcluster",
       x = "Cluster 5 Subcluster",
       y = "Percentage of Subcluster") +
    geom_text(aes(label = paste0(round(Percentage), "%")),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 3)  +
  theme_minimal()
```


```{r}
data <- as.data.frame(table(so.clust5$subclust_5.08, so.clust5$clonotype_frequency))
colnames(data) <- c("Subcluster", "Clone_Num", "Freq")

result <- data %>%
  group_by(Subcluster) %>%
  mutate(Percentage = Freq / sum(Freq) * 100) %>%
  ungroup()

result$Clone_Num <- unfactor(result$Clone_Num)

result$Clone_Num[result$Clone_Num == 1] <- "Unique Clonotype"

result$Clone_Num <- factor(result$Clone_Num, levels = c("Unique Clonotype", 2, 3, 4, 5))

ggplot(result, aes(x = factor(Subcluster), y = Percentage, fill = Clone_Num)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "Clone_Num/Clonal Expansion Distribution in Each Subcluster",
       x = "Cluster 5 Subcluster",
       y = "Percentage of Subcluster Cells with each Clone #",
       fill = "Cells with the Same Clonotype") +
      geom_text(aes(label = paste0(round(Percentage), "%")),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 3)  +
  theme_minimal()
```


```{r}
saveRDS(so.clust5, file = paste0(data_dir, "only_subclust5_12_08.rds"))
```



# Adding Subcluster to Seurat object with all cells

```{r}
#setwd(data_dir)
seurat_object_new_new <- readRDS(seurat_object_new_new, file = paste0(data_dir, "seurat_new_12_08.rds"))
```

```{r}
temp_subclust <- c()
temp_subclust$subclust <- so.clust5$subclust_5.08
names(temp_subclust$subclust) <- names(so.clust5$subclust_5.08)
temp_subclust$cell_barcode <- names(so.clust5$subclust_5.08)
#temp_subclust <- as.data.frame(temp_subclust)

seurat_object_new_new$subclust5 <- temp_subclust$subclust[match(so.clust5@meta.data[["cell_barcode"]], temp_subclust$cell_barcode)]
```


```{r}
plot.list <- lapply(X = ob.list, FUN = function(x) {
    DimPlot(x, reduction = "umap", label = FALSE, label.size = 4)
})

p1 <- DimPlot(seurat_object_new_new, reduction = "umap_rna", group.by = "subclust5", label = TRUE, split.by = "Mouse_New", ncol = 4)
#seurat_object_new_new$Infection
```

```{r}
seurat_object_new_new$subclust5[is.na(seurat_object_new_new$subclust5)] <- "Not in Cluster 5"
seurat_object_new_new$subclust5 <- as.factor(seurat_object_new_new$subclust5)
```

```{r}
saveRDS(seurat_object_new_new, file = paste0(data_dir, "so_allcells_subclust5_12_08.rds"))
```



###

# ONLY CD4 Diffusion Map
```{r}
AverageExpression(seurat_object, assay = "ADT", features = "CD4", group.by = "RNA_clusters")

DefaultAssay(seurat_object) <- "ADT"
FeaturePlot(seurat_object, features = "CD4", reduction = "umap_rna")
so_CD4 <- subset(x = seurat_object, subset = CD4 > 2.0)
so_CD4$RNA_clust_annot <- factor(so_CD4$RNA_clust_annot)

sce_CD4 <- as.SingleCellExperiment(so_CD4)
```

https://broadinstitute.github.io/2019_scWorkshop/functional-pseudotime-analysis.html#diffusion-map-pseudotime


```{r}
library(destiny)
dm_CD4 <- DiffusionMap(sce_CD4, verbose = TRUE)
```

```{r}
#cellLabels <- sce_CD4$RNA_clusters
#cellLabels <- sce_CD4$RNA_clust_annot
#cellLabels <- sce_CD4$Mouse
cellLabels <- sce_CD4$Organ
tmp <- data.frame(DC1 = eigenvectors(dm_CD4)[, 1],
                  DC2 = eigenvectors(dm_CD4)[, 2],
                  DC3 = eigenvectors(dm_CD4)[, 3],
                  DC4 = eigenvectors(dm_CD4)[, 4],
                  Samples = cellLabels)
#pdf("./DC1_DC2.pdf", w=11, h=8.5)
ggplot(tmp, aes(x = DC1, y = DC2, colour = Samples)) +
  geom_point()  + 
  xlab("Diffusion component 1") + 
  ylab("Diffusion component 2") +
  theme_classic()
```

```{r}
sce_CD4$pseudotime_diffusionmap <- rank(eigenvectors(dm_CD4)[,1])    # rank cells by their dpt
ggplot(as.data.frame(colData(sce_CD4)), 
       aes(x = pseudotime_diffusionmap, 
           y = RNA_clust_annot, colour = RNA_clust_annot)) +
    geom_point() +
    xlab("Diffusion component 1 (DC1)") + ylab("Timepoint") +
    ggtitle("Cells ordered by DC1")
```


```{r}
plotExpression(sce_CD4, "Il23r", x = "pseudotime_diffusionmap", 
               colour_by = "RNA_clust_annot", show_violin = FALSE,
               show_smooth = TRUE)
```


```{r}
# Can select which tips you would like to acknowledge
dpt <- DPT(dm_CD4)

grid.arrange(plot(dpt))
```



# HOW TO MAKE A GENE SET FEATURE PLOT
```{r}
gene.set <- sample(x = rownames(x = seurat_object@assays[["RNA"]]), size = 100, replace = FALSE)

Il17.gene.set <- c("Il17a", "Il23r", "Rorc")

# Get mean expression of genes of interest per cell
mean.exp <- colMeans(x = seurat_object@assays[["RNA"]][Il17.gene.set, ], na.rm = TRUE)

# Add mean expression values in 'object@meta.data$gene.set.score'
if (all(names(x = mean.exp) == rownames(x = seurat_object@meta.data))) {
  cat("Cell names order match in 'mean.exp' and 'object@meta.data':\n", 
      "adding gene set mean expression values in 'object@meta.data$gene.set.score'")
  seurat_object@meta.data$Il17.gene.set.score <- mean.exp
}

# Plot mean expression using Seurat::FeaturePlot()
FeaturePlot(object = seurat_object, features = "Il17.gene.set.score")
```









### OLD, NOT 12/11/2023
### OK





# Cluster 5 - Finding Markers between Candida and Non-Candida for each Organ

```{r}
organs <- c(unique(seurat_object$Organ))

for (i in organs){
  Idents(object = seurat_object) <- "RNA_clusters"
  DefaultAssay(seurat_object) <- "RNA"
  seurat_object.sub <- subset(x = seurat_object, subset = (RNA_clusters == 5))
  seurat_object.organ <- subset(x = seurat_object.sub, subset = (Organ == i))
  
  Idents(object = seurat_object.organ) <- "Infection"
  
  DefaultAssay(seurat_object.organ) <- "RNA"
  Candida_cluster5.markers_per_organ <- FindMarkers(seurat_object.organ, ident.1 = "Ca", only.pos = TRUE, test.use = "DESeq2", logfc.threshold = 0.05, min.pct = 0.05)
  
  write.csv(Candida_cluster4.markers_per_organ, paste(i, "_CA_RNA_Markers.csv", sep = ""))
}
```



# Increasing RNA clustering resolution to split Cluster 5
Here, I'm trying to re-run the analysis from the Commands section of the seurat object in order to change the resolution and check cluster stability:
```{r}
set.seed(42)
DefaultAssay(seurat_object.lung) <- "RNA"
DefaultAssay(seurat_object.colon) <- "RNA"
# seurat_object.lung, seurat_object.mlns, seurat_object.colon

seurat_object.lung <- NormalizeData(object = seurat_object.lung)
seurat_object.lung <- FindVariableFeatures(object = seurat_object.lung)
seurat_object.lung <- ScaleData(object = seurat_object.lung)
seurat_object.lung <- RunPCA(object = seurat_object.lung)
seurat_object.lung <- FindNeighbors(object = seurat_object.lung)
seurat_object.lung <- FindClusters(object = seurat_object.lung, resolution = c(2.5))

seurat_object.colon <- NormalizeData(object = seurat_object.colon)
seurat_object.colon <- FindVariableFeatures(object = seurat_object.colon)
seurat_object.colon <- ScaleData(object = seurat_object.colon)
seurat_object.colon <- RunPCA(object = seurat_object.colon)
seurat_object.colon <- FindNeighbors(object = seurat_object.colon)
seurat_object.colon <- FindClusters(object = seurat_object.colon, resolution = c(2.5))
```

# Cluster Similarity between RNA Clusters and Cluster 5 Split Clusters

Lung
```{r}
cluster_matrix <- as.matrix(table(seurat_object.lung$RNA_clusters, seurat_object.lung$RNA_snn_res.2.5))

rownames(cluster_matrix) <- paste("RNA", rownames(cluster_matrix))
colnames(cluster_matrix) <- paste("Split", colnames(cluster_matrix))

# Create heatmap
pheatmap(cluster_matrix, 
        #col = colorRampPalette(c("yellow", "blue"))(20),
        main = "Lung Cluster Similarity Heatmap between RNA and RNA 2.5 Resolution")
        #labels_row = make_bold_names(cluster_matrix, rownames, c("WNN 4")),
        #labels_col = make_bold_names(cluster_matrix, colnames, c("RNA 5")))
        #xlab = "RNA Clusters",
        #ylab = "WNN Clusters")
```

Based on this graph, RNA 5 is split into clusters 13, 16 and 17 for lung.

Colon
```{r}
cluster_matrix <- as.matrix(table(seurat_object.colon$RNA_clusters, seurat_object.colon$RNA_snn_res.2.5))

rownames(cluster_matrix) <- paste("RNA", rownames(cluster_matrix))
colnames(cluster_matrix) <- paste("Split", colnames(cluster_matrix))

# Create heatmap
pheatmap(cluster_matrix, 
        #col = colorRampPalette(c("yellow", "blue"))(20),
        main = "Colon Cluster Similarity Heatmap between RNA and RNA 2.5 Resolution")
        #labels_row = make_bold_names(cluster_matrix, rownames, c("WNN 4")),
        #labels_col = make_bold_names(cluster_matrix, colnames, c("RNA 5")))
        #xlab = "RNA Clusters",
        #ylab = "WNN Clusters")
```

Based on this graph, RNA 5 is split into clusters 13 and 14 for colon.

```{r}
Idents(object = seurat_object.lung) <- "RNA_snn_res.2.5"
Idents(object = seurat_object.colon) <- "RNA_snn_res.2.5"

DimPlot(seurat_object.lung, reduction = "umap_rna", label = TRUE, repel = TRUE) + ggtitle("Lung Split 5 Clusters")

DotPlot(seurat_object.lung, features = c(stem_like_gois, activation_like_gois)) + RotatedAxis() + ggtitle("Lung Split 5 Clusters by Stem and Activation Markers")

DimPlot(seurat_object.colon, reduction = "umap_rna", label = TRUE, repel = TRUE) + ggtitle("Colon Split 5 Clusters")

DotPlot(seurat_object.colon, features = c(stem_like_gois, activation_like_gois)) + RotatedAxis() + ggtitle("Colon Split 5 Clusters by Stem and Activation Markers")

#stem_like_gois <- c("Ccr7", "Il7r", "Sell", "Tcf7", "Tcf1", "Slamf6", "Klf3", "Il16ra")
#activation_like_gois <- c("Cd44", "Cd69", "Il17a", "Il17f", "Il22", "Ifng", "Mki67", "Csf2")
```
For the Lung, we are primarily focused on clusters 13, 16 and 17. 
13 has higher levels of Ccr7, Sell, Slamf6, which implies it's the stem-like cluster
16 has higher levels of Cd44 and some Il17a, which implies it's the activation-like cluster
17 has higher levels of Il17a and Il17f, which implies it's the activation-like cluster

For the Lung, we are primarily focused on clusters 13, and 14. 
13 has higher levels of Il17a, Cd44, and Il17f, which implies it's the activation-like cluster. But it has Il17r also at higher levels
14 doesn't really have higher expression levels in any gene of interest.

```{r}
cluster.5.split.markers.lung.stem <- FindMarkers(seurat_object.lung, ident.1 = 13, ident.2 = 16, only.pos = TRUE)
cluster.5.split.markers.lung.gen <- FindMarkers(seurat_object.lung, ident.1 = 16, ident.2 = 13, only.pos = TRUE)
cluster.5.split.markers.colon.stem <- FindMarkers(seurat_object.colon, ident.1 = 14, ident.2 = 13, only.pos = TRUE)
cluster.5.split.markers.colon.gen <- FindMarkers(seurat_object.colon, ident.1 = 13, ident.2 = 14, only.pos = TRUE)
```

# Genes of Interest when splitting cluster 5:
```{r}
cluster.5.split.markers.lung.stem[row.names(cluster.5.split.markers.lung.stem) %in% stem_like_gois,]
cluster.5.split.markers.lung.gen[row.names(cluster.5.split.markers.lung.gen) %in% activation_like_gois, ]

cluster.5.split.markers.colon.stem[row.names(cluster.5.split.markers.colon.stem) %in% stem_like_gois, ]
cluster.5.split.markers.colon.gen[row.names(cluster.5.split.markers.colon.gen) %in% activation_like_gois, ]
```

Could also make a seurat object with only cluster 5 cells instead


