---
title: "Candida_sc_CD4_diff"
author: "Nathan Morris"
date: "2024-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

### Data Import
```{r echo = TRUE, include= TRUE, error = FALSE, warning = FALSE, message=FALSE}
# Data Import
library(Seurat)
library(scater)
library(dplyr)
library(SingleR)
library(celldex)
library(scRNAseq)
library(scran)
library(gridExtra)
library(pheatmap)
library(ess)
library(reshape2)
library(ggpubr)
library(scCustomize)
library(gridExtra)
library(monocle3)

## For Reticulate (Python-R wrapper)
conda_list()
use_condaenv("r-reticulate")    #activate an environment e.g py2_env
py_install("opentsne")
reticulate::py_install("phate", pip=TRUE)
devtools::install_github("KrishnaswamyLab/phateR")
py_install("fa2")

data_dir <- "/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Candida_Migratory_TCR_Project/Data/"

setwd(data_dir)
```

```{r}
seurat_object <- readRDS(file = paste0(data_dir, "dataset_clean_scRNA_ADT_new_metadata.Rds"))
```

# 3. Subsetting only CD4 Differentiated Cells based on Markers

Question - How would we like to subset cells - 
1. Subset by first clustering them all, finding marker genes and then selecting certain clusters with marker genes?

2. Subset the cells by Expression of CD4 and Sell (Differentiated CD4 T cells), then re-cluster



```{r}
FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD8A")

FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "Sell")
FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD44")

FeatureScatter(seurat_object, feature1 = "CD44", feature2 = "Sell")
FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD62L")

Idents(seurat_object) <- "RNA_clusters"
RidgePlot(seurat_object, features = "CD44")
# This makes me think we can separate at like a 1.5 expression level in order to 


Idents(seurat_object) <- "RNA_clusters"
RidgePlot(seurat_object, features = "Sell")
```

```{r}
DefaultAssay(seurat_object) <- "ADT"
Idents(seurat_object) <- "RNA_clusters"

seurat_object_CD4 <- subset(x = seurat_object, subset = CD4 > 1.5)

DefaultAssay(seurat_object_CD4) <- "RNA"
seurat_object_CD4_diff <- subset(x = seurat_object_CD4, subset = Sell < 0.5)

seurat_object$CD4_cell <- seurat_object$cell_barcode %in% seurat_object_CD4$cell_barcode

seurat_object$CD4_cell_diff <- seurat_object$cell_barcode %in% seurat_object_CD4_diff$cell_barcode

DimPlot(seurat_object, group.by = "CD4_cell_diff", reduction = "umap_rna")

FeaturePlot(seurat_object, features = "CD62L", reduction = 'umap_rna')
```

```{r}
seurat_object_CD4_diff <- seurat_object_CD4_diff %>% NormalizeData(verbose = FALSE) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(verbose = TRUE)
seurat_object_CD4_diff <- seurat_object_CD4_diff %>% RunPCA(pc.genes = seurat_object_CD4_diff@var.genes, npcs = 30, verbose = FALSE) %>%  RunUMAP(dims = 1:25, reduction.name = "umap_cd4_") %>% RunTSNE(dims = 1:25, reduction.name = "tsne_cd4_") %>% FindNeighbors(dims = 1:25,compute.SNN=TRUE) %>% FindClusters(resolution = c(0.1,0.2,0.4,0.6,0.8,1), cluster.name = "cd4_") %>% DoopenTSNE(reduction_save = "openTSNE_cd4_",dims_use = 1:25) %>% DoForceAtlas2(reduction_use = "pca", reduction_save = "fa2_cd4_", dims_use = 1:25)
```

Or you can use this if you'd like to take the Annotated clusters
```{r}
seurat_object_CD4_diff <- readRDS(file = paste0(data_dir, "seurat_object_CD4_diff.rds"))
```


# 3.A Subsetted cells - Old RNA clusters in old RNA UMAP

```{r}
Idents(object = seurat_object_CD4_diff) <- "RNA_clusters"
#Idents(object = seurat_object_high_clust) <- "ADT_snn_res.1.8"
```

```{r}
DimPlot(seurat_object_CD4_diff, reduction = "umap_rna", label = FALSE, repel = TRUE, split.by = "Infection")
```

# 3.B Subsetted cells - Old RNA clusters in new Dimension Reduction
```{r}
DimPlot(seurat_object_CD4_diff, reduction = "pca", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4_diff, reduction = "umap_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4_diff, reduction = "tsne_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4_diff, reduction = "openTSNE_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4_diff, reduction = "fa2_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
```

# 3.C Subsetted cells - New RNA clusters in new Dimension Reduction

```{r}
Idents(object = seurat_object_CD4_diff) <- "RNA_snn_res.0.8"

DimPlot(seurat_object_CD4_diff, reduction = "umap_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4_diff, reduction = "tsne_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4_diff, reduction = "openTSNE_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4_diff, reduction = "fa2_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
```


```{r}
seurat_object_CD4_diff$CD4_diff_clust <- seurat_object_CD4_diff$RNA_snn_res.0.8
```

```{r}
DimPlot(seurat_object_CD4_diff, reduction = "umap_cd4_", label = FALSE, repel = TRUE, split.by = "Mouse_New", group.by = "CD4_diff_clust")
```

```{r}
df_cd4_mouse <- table(seurat_object_CD4_diff$Mouse_New, seurat_object_CD4_diff$CD4_diff_clust, seurat_object_CD4_diff$Organ)

temp <- as.data.frame(melt(df_cd4_mouse))
colnames(temp) <- c("Mouse", "Cluster","Organ", "Freq")
```


```{r}
ggplot(temp, aes(x = Mouse, y = Freq, color = Mouse)) +
  geom_boxplot() +
  labs(
    x = "Cluster",
    y = "Frequency",
    title = "Boxplot of Frequency by Cluster and Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(Organ ~ Cluster) +
  theme(legend.position = "top")
```

# 3.C Il-17 Gene Set Score

Below gene sets are from: https://www.nature.com/articles/s41590-020-00836-7 / https://static-content.springer.com/esm/art%3A10.1038%2Fs41590-020-00836-7/MediaObjects/41590_2020_836_MOESM1_ESM.pdf

```{r}
Il17.gene.set <- c("Il17a", "Il17f", "Il22", "Ccr6", "Il23r", "Rorc", "Tmem176a")
Th1.gene.set <- c("Ifng", "Tbx21", "Cxcr3", "Nkg7", "Fasl", "Il2", "Tnf", "Il12rb1", "Il12rb2", "Gzmb")
Th2.gene.set <- c("Gata3", "Il4", "Il5", "Il13", "Il10", "Areg", "Maf", "Il17rb")
Tfh.gene.set <- c("Cxcr5", "Izumo1r", "Nt5e", "Bcl6", "Slamf6", "Cd200", "Ccr7", "Tcf7")

# Get mean expression of genes of interest per cell
mean.exp <- colMeans(x = seurat_object_CD4_diff@assays[["RNA"]][Il17.gene.set, ], na.rm = TRUE)

# Add mean expression values in 'object@meta.data$gene.set.score'
if (all(names(x = mean.exp) == rownames(x = seurat_object_CD4_diff@meta.data))) {
  cat("Cell names order match in 'mean.exp' and 'object@meta.data':\n", 
      "adding gene set mean expression values in 'object@meta.data$gene.set.score'")
  seurat_object_CD4_diff@meta.data$Il17.gene.set.score <- mean.exp
}
```

# 3.C-1 Il17 Score in Old UMAP

```{r}
# Plot mean expression using Seurat::FeaturePlot()
FeaturePlot(object = seurat_object_CD4_diff, features = "Il17.gene.set.score", reduction = "umap_rna") 
#scale_color_manual(values = c("grey", "forestgreen", "yellow", "red")) + labs(color = "Expression")

FeaturePlot(object = seurat_object_CD4_diff, features = "Il17.gene.set.score", reduction = "umap_cd4_", split.by = "Infection") 
```

# 3.C-2 Il17 Score in New Dimension Reductions

```{r}
FeaturePlot(object = seurat_object_CD4_diff, features = "Il17.gene.set.score", reduction = "umap_cd4_")

FeaturePlot(object = seurat_object_CD4_diff, features = "Il17.gene.set.score", reduction = "tsne_cd4_")

FeaturePlot(object = seurat_object_CD4_diff, features = "Il17.gene.set.score", reduction = "openTSNE_cd4_")

FeaturePlot(object = seurat_object_CD4_diff, features = "Il17.gene.set.score", reduction = "fa2_cd4_")
```

```{r}
saveRDS(seurat_object_CD4_diff, file = paste0(data_dir, "seurat_object_CD4_diff.rds"))
```


# 3.D Selecting Clusters that look Differentiated on CD44/CD4

```{r}
seurat_object <- readRDS(file = paste0(data_dir, "dataset_clean_scRNA_ADT_new_metadata.Rds"))
```

```{r}
Idents(object = seurat_object) <- "RNA_clusters"
subselected_so <- subset(x = seurat_object, idents = c("7", "9", "5", "13", "11", "12"))

DimPlot(subselected_so, reduction = "umap_rna", group.by = "RNA_clusters", split.by = "Infection")
```