---
title: "Candida_tcr_duplicate"
author: "Nathan Morris"
date: "2023-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

# Directory Set Up

```{r}
data_dir <- "/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Candida_Migratory_TCR_Project/Data/"

setwd(data_dir)
```

Here, we use the pre-existing seurat object.
```{r}
seurat_object <- readRDS(file = paste0(data_dir, "/dataset_w_annot.rds"))

#seurat_object_WNN <- readRDS("Candida_ds_clean_WNN.rds") # I didn't include WNN analysis in this document
#seurat_object <- readRDS(file = paste0(data_dir, "dataset_post_sc_Analysis.rds")) #If you want without WNN Clusters and without TCR Information

seurat_object_pseudo <- readRDS(file = paste0(data_dir, "pseudotime_seurat.rds"))
```

And we get the object containing our filtered TCR data
```{r}
IGT27_TCR <- read.table(paste0(data_dir,"IGT27_for_analysis.tsv"), header = TRUE, sep = "\t")
```

# Load Packages

```{r echo = TRUE, include= TRUE, error = FALSE, warning = FALSE, message=FALSE}
library(ggplot2)
library(Seurat)
library(tibble)
library(dplyr)
library(pheatmap)
library(tidyr)
library(usedist)
library(ggvenn)
library(stringdist)
library(patchwork)
library(djvdj)
library(data.table)
```

# 2. TCR Duplicates not specifying Organ

Here, we are checking for clonotypes that are found in multiple cells, but aren't specifically found in multiple organs. Therefore, this means that they proliferated but weren't migratory.

## 2.A Duplicate Column Creation

```
duplicate_clonotypes <- IGT27_TCR %>%
  group_by(clonotype, Mouse) %>%
  dplyr::filter(n_distinct(Barcode) > 1)

IGT27_TCR$duplicate <- IGT27_TCR$Barcode %in% duplicate_clonotypes$Barcode
```

## 2.B Duplicate Clonotype Cell Count by Mouse/Sample

Now we want to check for proliferation. Therefore, for each duplicated clonotype, we want to how many cells have this clonotype. We also want to check between mice, to see if the infection led to more proliferation of a certain clonotype.

Separating by Mouse, checking the Organ distribution, whether they are paired or not, WNN Clusters and what Organ the non-paired but duplicated clonotypes are in.

```{r}
duplicate_df_list <- list()
duplicate_df_plots <- list()
duplicate_df_plots_only <- list()

mice <- c("mouse0082", "mouse0080", "mouse0081")

mouse_dict <- c(
  "mouse0082" = "No Candida Male",
  "mouse0081" = "Candida Female",
  "mouse0080" = "Candida Male"
)

for (mouse in mice){
  mouse_desc <- mouse_dict[mouse]
  duplicate_df_list[[paste0(mouse, "_duplicate_freq")]] <- as.data.frame(table(subset(IGT27_TCR, Mouse == mouse & duplicate == "TRUE")$clonotype))
  
  p1 <- ggplot(duplicate_df_list[[paste0(mouse, "_duplicate_freq")]], aes(x = Freq)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 10) +
  labs(x = "Number of Cells with the same Clonotype", y = "Clonotype Count") +
  ggtitle(paste0(mouse_desc, " Duplicate TCR Frequencies")) +
  theme_minimal()
  
  duplicate_df_list[[paste0(mouse, "_not_duplicate_freq")]] <- as.data.frame(table(subset(IGT27_TCR, Mouse == mouse & duplicate == "FALSE")$clonotype))
  
  p2 <- ggplot(duplicate_df_list[[paste0(mouse, "_not_duplicate_freq")]], aes(x = Freq)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 10) +
  labs(x = "Number of Cells with the same Clonotype", y = "Clonotype Count") +
  ggtitle(paste0(mouse_desc, " Not Duplicate TCR Frequencies")) +
  theme_minimal()
  
  duplicate_df_plots[[mouse]] <- p1 + p2 
  duplicate_df_plots_only[[mouse]] <- p1
}
```

Here, I show how many clonotypes are shared between a certain number of cells.

```{r}
duplicate_df_plots_only[["mouse0082"]]
duplicate_df_plots_only[["mouse0081"]]
duplicate_df_plots_only[["mouse0080"]]
```

From the above you can see that when looking at cells with proliferating TCR clonotypes (duplicates) the non-infected male has a maximum number of cells with the same clonotype, which is 14. The infected female has a maximum number of cells with the same clonotype of 6. 

Therefore, the non-infected male shows higher proliferation of a single TCR clonotype than the infected female or infected male.

## 2.C Duplicate Clonotype Cell Count Total (All Mice)

```{r}
TCR_TRUE <- as.data.frame(table(subset(IGT27_TCR, duplicate == "TRUE")$clonotype))
TCR_FALSE <- as.data.frame(table(subset(IGT27_TCR, duplicate == "FALSE")$clonotype))

p7 <- ggplot(TCR_TRUE, aes(x = Freq)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 10) +
  labs(x = "Number of Cells with the same Clonotype", y = "Clonotype Count") +
  ggtitle("All Mice Duplicate TCR Frequencies") +
  theme_minimal()

p8 <- ggplot(TCR_FALSE, aes(x = Freq)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 10) +
  labs(x = "Number of Cells with the same Clonotype", y = "Clonotype Count") +
  ggtitle("All Mice Not-Duplicate TCR Frequencies") +
  theme_minimal()

p7 + p8
```

It it shown again that the maximum proliferation of a single TCR sequence is 14 cells. These 14 cells are found in the Non-infected Male. 

Max 6 cells are found with the same TCR clonotype for the Infected mice (found in Infected Female.)

## 2.D Comparing Infected vs Non-infected Duplicate TCR Clonotypes

Here, we check for shared paired clonotypes between the non-infected and the infected mice.

```{r}
Candida_Duplicate_Clonotypes <- c(duplicate_df_list$mouse0081_duplicate_freq$Var1, duplicate_df_list$mouse0080_duplicate_freq$Var1)
No_Candida_Duplicate_Clonotypes <- duplicate_df_list$mouse0082_duplicate_freq$Var1

sum((Candida_Duplicate_Clonotypes %in% No_Candida_Duplicate_Clonotypes))
sum((No_Candida_Duplicate_Clonotypes %in% Candida_Duplicate_Clonotypes))
```

There are no clonotypes that are shared between the Infected and the Non-infected samples. We would expect there to be a clonotype in the non-infected which has proliferated in the infected, but it is not the case.

## 2.E RNA Cluster, Organ and Mouse Distribution

```{r}
dataframes_duplicates <- list()
parameters <- c("RNA_clusters", "Organ", "Mouse")
for (parameter in parameters){
  tmp_df_inf <- as.data.frame(table(IGT27_TCR[IGT27_TCR$duplicate == TRUE & IGT27_TCR$Infection == "Yes" ,][,parameter]))
  tmp_df_noinf <- as.data.frame(table(IGT27_TCR[IGT27_TCR$duplicate == TRUE & IGT27_TCR$Infection == "No" ,][,parameter]))
  merged_df <- full_join(tmp_df_inf, tmp_df_noinf, by = "Var1") %>% 
  replace_na(list(Freq.x = 0, Freq.y = 0))
  #merged_df <- merge(tmp_df_inf, tmp_df_noinf, by = "Var1", suffixes = c("_Infection", "_NoInfection"))
  colnames(merged_df) <- c(parameter, "Freq_Infection", "Freq_NoInfection")
  dataframes_duplicates[[parameter]] <- merged_df
}

df_long <- pivot_longer(dataframes_duplicates$RNA_clusters, cols = starts_with("Freq"), names_to = "Infection_Status", values_to = "Frequency")
```

```{r}
ggplot(df_long, aes(x = RNA_clusters, y = Frequency, fill = Infection_Status)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "RNA Cluster", y = "Number of Cells") +
  ggtitle("Number of Cells with Duplicate Clonotypes *not specifying Organ* in RNA Clusters") +
  theme_minimal()
```

```{r}
table(IGT27_TCR$duplicate, IGT27_TCR$RNA_clusters, IGT27_TCR$Mouse_New)
#table(IGT27_TCR$paired_across_organ, IGT27_TCR$RNA_clusters, IGT27_TCR$Mouse_New)
```

This is kind of interesting, shows the distribution of the duplicate cells (proliferating, as https://www.tandfonline.com/doi/full/10.1080/2162402X.2020.1737369 says) is different in each Mouse


## 2.F - Clonotype Organ Split

```{r}
max_clonotypes <- TCR_TRUE[TCR_TRUE$Freq == max(TCR_TRUE$Freq),]$Var1

organ_split_max_clonotype <- matrix(NA, nrow = length(max_clonotypes), ncol = length(unique(IGT27_TCR$Organ)))
colnames(organ_split_max_clonotype) <- c(unique(IGT27_TCR$Organ)[1:3], "Mouse", "Infection")
rownames(organ_split_max_clonotype) <- max_clonotypes


for (i in seq(1:length(max_clonotypes))){
   temp_df <- table(IGT27_TCR$Organ[IGT27_TCR$clonotype == max_clonotypes[i]]) 
   organ_split_max_clonotype[i, "Colon"] <- temp_df["Colon"]
   organ_split_max_clonotype[i, "mLns"] <- temp_df["mLns"]
   organ_split_max_clonotype[i, "Lung"] <- temp_df["Lung"]
   organ_split_max_clonotype[i, "Mouse"] <- IGT27_TCR[IGT27_TCR$clonotype == max_clonotypes[i],]$Mouse[1]
   organ_split_max_clonotype[i, "Infection"] <- IGT27_TCR[IGT27_TCR$clonotype == max_clonotypes[i],]$Infection[1]
}
```

```{r}
organ_split_max_clonotype
```

### Maximum for Each Mouse
```{r}
# Mouse 0 - Candida infected Male
mouse0_max <- duplicate_df_list$mouse0080_duplicate_freq[duplicate_df_list$mouse0080_duplicate_freq$Freq == max(duplicate_df_list$mouse0080_duplicate_freq$Freq),]$Var1
paste("The clonotype with maximum cells for the Infected Male: ", mouse0_max)
table(IGT27_TCR$Organ[IGT27_TCR$clonotype == mouse0_max])


# Mouse 1 - Candida Infected Female
mouse1_max <- duplicate_df_list$mouse0081_duplicate_freq[duplicate_df_list$mouse0081_duplicate_freq$Freq == max(duplicate_df_list$mouse0081_duplicate_freq$Freq),]$Var1
paste("The clonotype with maximum cells for the Infected Female: ", mouse1_max)
table(IGT27_TCR$Organ[IGT27_TCR$clonotype == mouse1_max])

# Mouse 2 - Non-infected Female
mouse2_max <- duplicate_df_list$mouse0082_duplicate_freq[duplicate_df_list$mouse0082_duplicate_freq$Freq == max(duplicate_df_list$mouse0082_duplicate_freq$Freq),]$Var1
paste("The clonotype with maximum cells for the Non-infected Female: ", mouse2_max)
table(IGT27_TCR$Organ[IGT27_TCR$clonotype == mouse2_max])

#table(IGT27_TCR[IGT27_TCR$clonotype == mouse2_max,]$RNA_clusters)
# Confirms that only 10 of these cells have RNA data
```

As you can see the clonotype with maximum cells is higher in the Non-infected mouse rather than the infected.

## 2.G Duplicate Statistics for Comparison



```
compare_df["# Cells", "Duplicates No Organ Specification"] <- sum(IGT27_TCR$duplicate == TRUE)

compare_df["% Cells", "Duplicates No Organ Specification"] <- sum(IGT27_TCR$duplicate == TRUE) / nrow(IGT27_TCR) * 100

compare_df["# Clonotypes", "Duplicates No Organ Specification"] <- length(unique(IGT27_TCR[IGT27_TCR$duplicate == TRUE,]$clonotype))

compare_df["% Clonotypes", "Duplicates No Organ Specification"] <- length(unique(IGT27_TCR[IGT27_TCR$duplicate == TRUE,]$clonotype)) / length(unique(IGT27_TCR$clonotype)) * 100

compare_df["Infected Cells #", "Duplicates No Organ Specification"] <- nrow(IGT27_TCR[IGT27_TCR$duplicate == TRUE & IGT27_TCR$Infection == "Yes",])

compare_df["Non-Infected Cells #", "Duplicates No Organ Specification"] <- nrow(IGT27_TCR[IGT27_TCR$duplicate == TRUE & IGT27_TCR$Infection == "No",])

compare_df["Infected Cells %", "Duplicates No Organ Specification"] <- nrow(IGT27_TCR[IGT27_TCR$duplicate == TRUE & IGT27_TCR$Infection == "Yes",]) / nrow(IGT27_TCR[IGT27_TCR$duplicate == TRUE,]) * 100

compare_df["% Paired Across Organ", "Duplicates No Organ Specification"] <- sum(IGT27_TCR[IGT27_TCR$duplicate == TRUE,]$paired_across_organ) / sum(IGT27_TCR$duplicate == TRUE) * 100
```

# 3. Comparing Paired and Duplicates

## 3.A Duplicate and Paired Clonotype Comparison Graphs

```{r}
dataframes_dup_intense_list <- list()

for (mouse in mice){
  tmp_df2 <- as.data.frame(IGT27_TCR[IGT27_TCR$duplicate == TRUE & IGT27_TCR$Mouse == mouse,])
  new_list <- list()
  for (row in seq(1:nrow(tmp_df2))){
    clonotype <- tmp_df2[row, ]$clonotype
    #print(clonotype)
    #print(paste(tmp_df2[tmp_df2$clonotype == clonotype, ]$RNA_clusters, collapse = ", "))
    n_clusters_per_clonotype <- length(unique(na.omit(tmp_df2[tmp_df2$clonotype == clonotype, ]$RNA_clusters)))
    n_organs_per_clonotype <- length(unique(tmp_df2[tmp_df2$clonotype == clonotype, ]$Organ))
    tmp_df2[row, "clusters_per_clonotype"] <- paste(tmp_df2[tmp_df2$clonotype == clonotype, ]$RNA_clusters, collapse = ", ")
    tmp_df2[row, "n_clusters_per_clonotype"] <- n_clusters_per_clonotype
    tmp_df2[row, "organs_per_clonotype"] <- paste(tmp_df2[tmp_df2$clonotype == clonotype, ]$Organ, collapse = ", ")
    tmp_df2[row, "n_organs_per_clonotype"] <- n_organs_per_clonotype
    #print(as.numeric(clusters_per_clonotype))
    #IGT27_TCR[IGT27_TCR$paired_across_organ == TRUE & IGT27_TCR$Mouse == "mouse0082" & IGT27_TCR$clonotype =="TRAV13N-1.TRAJ6.CAMREMGGNYKPTF.TRBV15.TRBJ1-1.CASSLGGSTEVFF",]
    #name(new_list[row]) <- tmp_clonotype
  }
  tmp_df2$n_clusters_per_clonotype <- as.factor(na.omit(tmp_df2$n_clusters_per_clonotype))
  tmp_df2$n_organs_per_clonotype <- as.factor(tmp_df2$n_organs_per_clonotype)
  dataframes_dup_intense_list[[mouse]] <- tmp_df2
}
```

The below data includes both the paired clonotype cells (paired across organ, same clonotype in multiple organs) and the duplicate clonotype cells (multiple cells from the same organ).

Here I created four graphs, the 

```{r}
df <- dataframes_dup_intense_list$mouse0082
df <- rbind(df, dataframes_dup_intense_list$mouse0080)
df <- rbind(df, dataframes_dup_intense_list$mouse0081)

df <- df %>%
  mutate(Mouse_New = ifelse(Mouse == "mouse0080", "Candida Infected Male", ifelse(Mouse == "mouse0081", "Candida Infected Female", ifelse(Mouse == "mouse0082", "No Infection Male", Mouse))))

level_order <- names(rev(sort(table(df$RNA_clusters, useNA = "ifany"))))
  
ggplot(df, aes(x= factor(RNA_clusters, level = level_order) , fill= n_clusters_per_clonotype)) + 
  geom_bar(aes(fill = n_clusters_per_clonotype)) +
  theme(plot.title = element_text(size = 7, face = "bold")) +
  labs(x = "RNA Clusters", y = "Number of Cells", fill = "Number of Clusters containing each Clonotype") +
  facet_grid(. ~ Infection, scales = "free_x", space = "free_x") + # Remove the fill legend
  ggtitle("Frequency of Cells with Duplicate/Paired Clonotypes in RNA Clusters by Infection and Cell Type (Multiple Clusters")


ggplot(df, aes(x= factor(RNA_clusters, level = level_order) , fill= n_clusters_per_clonotype)) + 
  geom_bar(aes(fill = n_clusters_per_clonotype)) +
  theme(plot.title = element_text(size = 7, face = "bold")) +
  labs(x = "RNA Clusters", y = "Number of Cells", fill = "Number of Clusters containing each Clonotype") +
  facet_grid(. ~ Mouse_New, scales = "free_x", space = "free_x") +  # Remove the fill legend
  ggtitle("Frequency of Cells with Duplicate/Paired Clonotypes in RNA Clusters by Mouse and Cell Type (Multiple Clusters)")

ggplot(df, aes(x= factor(RNA_clusters, level = level_order), fill= n_organs_per_clonotype)) + 
  geom_bar(aes(fill = n_organs_per_clonotype)) +
  theme(plot.title = element_text(size = 7, face = "bold")) +
  labs(x = "RNA Clusters", y = "Number of Cells", , fill = "Number of Organs containing each Clonotype") +
  facet_grid(. ~ Infection, scales = "free_x", space = "free_x") +  # Remove the fill legend
  ggtitle("Frequency of Cells with Duplicate/Paired Clonotypes in RNA Clusters by Infection and Organ")

ggplot(df, aes(x= factor(RNA_clusters, level = level_order), fill= n_organs_per_clonotype)) + 
  geom_bar(aes(fill = n_organs_per_clonotype)) +
  theme(plot.title = element_text(size = 7, face = "bold")) +
  labs(x = "RNA Clusters", y = "Number of Cells", , fill = "Number of Organs containing each Clonotype") +
  facet_grid(. ~ Mouse_New, scales = "free_x", space = "free_x") +  # Remove the fill legend
  ggtitle("Frequency of Cells with Duplicate/Paired Clonotypes in RNA Clusters by Mouse and Organ")
```

## NEW 09/28
They liked this graph, but didn't like how uninsightful it is when comparing the different organs to not have the different organs marked.

```{r}
ggplot(df, aes(x= factor(RNA_clusters, level = level_order), fill= n_organs_per_clonotype)) + 
  geom_bar(aes(fill = n_organs_per_clonotype)) +
  theme(plot.title = element_text(size = 7, face = "bold")) +
  labs(x = "RNA Clusters", y = "Number of Cells", , fill = "Num of Organs containing each /n Clonotype") +
  facet_grid(rows = vars(Organ), cols = vars(Mouse_New), scales = "free_x", space = "free_x") +  # Remove the fill legend
  ggtitle("Frequency of Cells with Duplicate/Paired Clonotypes in RNA Clusters by Infection and Organ")
```

######### 10/03/2023 Break

## 3.B Additional Comparison Analysis

```
compare_RNA_df_long <- tidyr::gather(merge(dataframes_duplicates$RNA_clusters, dataframes_paired$RNA_clusters, by='RNA_clusters'), key = "FrequencyType", value = "Frequency", -RNA_clusters)
compare_organ_df_long <- tidyr::gather(merge(dataframes_duplicates$Organ, dataframes_paired$Organ, by='Organ'), key = "FrequencyType", value = "Frequency", -Organ)
compare_mouse_df_long <- tidyr::gather(merge(dataframes_duplicates$Mouse, dataframes_paired$Mouse, by='Mouse'), key = "FrequencyType", value = "Frequency", -Mouse)
```

```
ggplot(compare_RNA_df_long, aes(x = RNA_clusters, y = Frequency, fill = FrequencyType)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "RNA_Cluster", y = "Number of Cells", fill = "Frequency Type") +
  theme_minimal()

ggplot(compare_organ_df_long, aes(x = Organ, y = Frequency, fill = FrequencyType)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Organ", y = "Number of Cells", fill = "Frequency Type") +
  theme_minimal()

ggplot(compare_mouse_df_long, aes(x = Mouse, y = Frequency, fill = FrequencyType)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Mouse", y = "Number of Cells", fill = "Frequency Type") +
  theme_minimal()
```

```
compare_df
```

# Using the mcPAS-TCR database

http://friedmanlab.weizmann.ac.il/McPAS-TCR/

```
mcPAS <- read.table(paste0(data_dir,"/mcPAS-TCR/McPAS-TCR.csv"), header = TRUE, sep = ",")

mcPAS_nona <- mcPAS[!is.na(mcPAS$CDR3.alpha.aa) & !is.na(mcPAS$CDR3.beta.aa),]

mcPAS_nona$CDR3 <- paste0(mcPAS_nona$CDR3.alpha.aa, mcPAS_nona$CDR3.beta.aa)

IGT27_TCR_try <- IGT27_TCR

IGT27_TCR_try$CDR3_all <- paste0(IGT27_TCR_try$AA.JUNCTION_alpha, IGT27_TCR_try$AA.JUNCTION_beta)

  
cdr3_matching <- function(list1, list2) {
  matched_pairs <- list()
  for (cdr3_1 in list1) {
    for (cdr3_2 in list2) {
      if (abs(nchar(cdr3_1) - nchar(cdr3_2)) <= 1) {  # Check if the length difference is at most 1
        mismatch_count <- sum(strsplit(cdr3_1, "")[[1]] != strsplit(cdr3_2, "")[[1]])
        if (mismatch_count <= 1) {
          matched_pairs <- c(matched_pairs, list(cdr3_1, cdr3_2))
        }
      }
    }
  }
  return(matched_pairs)
}

igt27list <- list(IGT27_TCR_try$CDR3_all)
mcpaslist <- list(mcPAS_nona$CDR3)

cdr3_matching(IGT27_TCR_try$igt27list, mcpaslist)


average_length <- function(string) {
  lengths <- nchar(string)  # Get the lengths of each string
  avg_length <- mean(lengths)  # Calculate the average length
  return(avg_length)
}


average_lengths_IGT27 <- sapply(IGT27_TCR_try$CDR3_all, average_length)
average_lengths_mcPAS <- sapply(list(mcPAS_nona$CDR3), average_length)

```

```{r}
cdr3_list <- cat(IGT27_TCR$AA.JUNCTION_beta, sep = "\n")

file_path <- "/cdr3_output.txt"

# Write the list elements to a text file
writeLines(cdr3_list, paste0(data_dir,file_path))

write.table(noquote(IGT27_TCR$AA.JUNCTION_beta), file = paste0(data_dir,file_path), sep = "\n",
            row.names = FALSE, col.names = FALSE, quote = FALSE)
```


