---
title: "Candida_sc_CD4_8"
author: "Nathan Morris"
date: "2024-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```

### Data Import
```{r echo = TRUE, include= TRUE, error = FALSE, warning = FALSE, message=FALSE}
# Data Import
library(Seurat)
library(scater)
library(dplyr)
library(SingleR)
library(celldex)
library(scRNAseq)
library(scran)
library(gridExtra)
library(pheatmap)
library(ess)
library(reshape2)
library(ggpubr)
library(scCustomize)
library(gridExtra)
library(monocle3)
library(cowplot)
library(reshape2)
library(reticulate)
library(ReductionWrappers)
library(paletteer)
library(paletteer)
library(cowplot)
library(SCpubr)
library(ggrepel)


data_dir <- "/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Candida_Migratory_TCR_Project/Data/"

setwd(data_dir)
```


```{r}
## For Reticulate (Python-R wrapper)
conda_list()
use_condaenv("r-reticulate")    #activate an environment e.g py2_env
py_install("opentsne")
reticulate::py_install("phate", pip=TRUE)
devtools::install_github("KrishnaswamyLab/phateR")
py_install("fa2")
```

```{r}
seurat_object <- readRDS(file = paste0(data_dir,"/dataset_post_QC.rds"))
```

# CD4

# 1.A Visualizing the splitting on CD4
```{r}
FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD8A") + geom_vline(xintercept = 0.75, linetype = "dashed", color = "red", size = 1.5) + geom_hline(yintercept = 0.75, linetype = "dashed", color = "blue", size = 1.5)

#FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD44") + geom_vline(xintercept = 0.75, linetype = "dashed", color = "red", size = 1.5)

Idents(seurat_object) <- "RNA_clusters"
RidgePlot(seurat_object, features = "CD4") + geom_vline(xintercept = 0.75, linetype = "dashed", color = "red", size = 1.5)
#FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "Sell")
#FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD44")
#FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD62L")
```

# 1.B Splitting into CD4 Seurat Object

```
DefaultAssay(seurat_object) <- "ADT"
Idents(seurat_object) <- "RNA_clusters"

seurat_object_CD4 <- subset(x = seurat_object, subset = CD4 > 0.75)

#saveRDS(seurat_object_CD4, file = paste0(data_dir, "/CD4_data/seurat_object_CD4.rds"))
```

```{r}
seurat_object_CD4 <- readRDS(file = paste0(data_dir, "/CD4_data/seurat_object_CD4.rds"))
```


# 1.C Dimension Reduction and Normalization

```
DefaultAssay(seurat_object_CD4) <- "RNA"

seurat_object_CD4 <- seurat_object_CD4 %>% NormalizeData(verbose = FALSE) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(verbose = TRUE)
seurat_object_CD4 <- seurat_object_CD4 %>% RunPCA(pc.genes = seurat_object_CD4@var.genes, npcs = 50, verbose = FALSE) %>%  RunUMAP(dims = 1:15, reduction.name = "umap_cd4_") %>% RunTSNE(dims = 1:15, reduction.name = "tsne_cd4_") %>% FindNeighbors(dims = 1:15,compute.SNN=TRUE) %>% FindClusters(resolution = c(0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2.0), cluster.name = "cd4_") %>% DoopenTSNE(reduction_save = "openTSNE_cd4_",dims_use = 1:15) %>% DoForceAtlas2(reduction_use = "pca", reduction_save = "fa2_cd4_", dims_use = 1:15)
```

```{r}
ElbowPlot(seurat_object_CD4, ndims = 50, reduction = "pca") + ggtitle("CD4 PCA Elbow Plot")
```


# 1.D - Visualization New CD4 clusters in new Dimension Reduction

```{r}
CD4_colors <- c("#373634", "#F7B6D2", "#F8EED1", "#14FFB1", "purple" ,paletteer_d("ggthemes::calc"))
```

Split by Infection
```{r}
Idents(object = seurat_object_CD4) <- "RNA_snn_res.1.2"
DimPlot(seurat_object_CD4, reduction = "pca", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4, reduction = "umap_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4, reduction = "tsne_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4, reduction = "openTSNE_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD4, reduction = "fa2_cd4_", label = FALSE, repel = TRUE, split.by = "Infection")
```

CD4 cluster metadata set based on a resolution of 1.2
```
seurat_object_CD4$CD4_clust <- seurat_object_CD4$RNA_snn_res.1.2
```

# 1.F Visualize Organ/Inf Split
```{r}
#Visualize Split by Mouse
Idents(object = seurat_object_CD4) <- "CD4_clust"
DimPlot(seurat_object_CD4, reduction = "tsne_cd4_", label = TRUE, repel = TRUE, split.by = "Mouse_New", group.by = "CD4_clust")

#Visualize Split by Organ and Infection
seurat_object_CD4$Inf_Organ <- paste0(seurat_object_CD4$Infection, "_", seurat_object_CD4$Organ)

DimPlot(seurat_object_CD4, reduction = "tsne_cd4_", label = FALSE, repel = TRUE, split.by = "Inf_Organ", ncol = 3)
```

# 1.G Cluster-Organ Frequency FIGURE 3 P1

```{r}
df_cd4_mouse <- table(seurat_object_CD4$Mouse_New, seurat_object_CD4$CD4_clust, seurat_object_CD4$Organ)
df_cd4_organ <- table(seurat_object_CD4$CD4_clust, seurat_object_CD4$Organ)

temp <- as.data.frame(melt(df_cd4_organ))
colnames(temp) <- c("Cluster","Organ", "Freq")

temp <- temp %>%
  group_by(Cluster) %>%
  mutate(Organ_Cluster_perc = Freq / sum(Freq) * 100)

#Fix the factor for visualization
temp$Cluster <- factor(temp$Cluster, levels = c("1", "9", "13", "7", "16", "14", "10", "5", "15", "2", "8", "6", "0", "4", "12", "11", "3"))

temp$Total_cells <- sum(temp$Freq)

temp <- temp %>%
  group_by(Cluster) %>% mutate(Cluster_perc = sum(Freq) / Total_cells * 100)

```

```{r}
organ_perc1_CD4 <- ggplot(temp, aes(x = Cluster, y = Organ_Cluster_perc, fill = Organ)) +
    geom_bar(stat = "identity", position = "fill") + 
    labs(title = "CD4 Clusters - Organ Proportion",
         x = "CD4 Cluster",
         y = "Proportion of each Cluster",
         fill = "Organ") + #theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x=element_blank(), legend.position = "none")  +   geom_label_repel(aes(label = sprintf("%.1f%%", Organ_Cluster_perc)), direction = ("y"),
               position = position_fill(vjust = 0.5), size = 3)
```

```{r}
cluster_sum <- aggregate(Cluster_perc ~ Cluster, data = temp, sum)

organ_perc2_CD4 <- ggplot(temp, aes(x = Cluster, y = Freq, fill = Organ)) +
    geom_bar(stat = "identity") + 
    #geom_text(data = Cluster_perc, aes(x = Cluster, y = max(temp$Freq) + 50, label = paste0(round(Cluster_perc, 1), "%")),
         #     position = position_dodge(width = 0.9), color = "black", size = 3) +
    labs(title = "CD4 Clusters - Total Number by Organ",
         x = "CD4 Cluster",
         y = "Number of Cells",
         fill = "Organ") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
```

```{r}
blank <- ggplot() + theme_void()

organ_perc3_CD4 <- DimPlot(seurat_object_CD4, group.by = "Organ", reduction = "tsne_cd4_") + ggtitle("CD4 TSNE by Organ") + NoLegend()

cd4_organ <- cowplot::plot_grid(organ_perc1_CD4,organ_perc2_CD4,organ_perc3_CD4,  ncol = 1, labels = c("A", "B", "C"))
```

# 1.H Visualizing Expression of Cluster-specific markers

```{r}
adt_specific_markers <- c("CD4", "CD8A", "TCRB", "TCRGD", "CD44", "CD62L")
rna_specific_markers <- c("Ccr7", "Foxp3", "Il17a", "Csf2", "Ifng", "Gzmb", "Tbx21", "Gata3", "Il5", "Il4") # Removes Il3 because level is 0 for all cells

# Function to create FeaturePlot for a specific marker
create_feature_plot <- function(marker) {
  FeaturePlot(seurat_object_CD4, features = marker, reduction = "tsne_cd4_", cols = c("yellow", "red")) + NoLegend() + theme(
    plot.title = element_text(size = 7, face = "bold"),
    axis.title.x = element_blank(),  # Remove x-axis label
    axis.title.y = element_blank()   # Remove y-axis label
  )
}

# Use lapply to create a list of plots for each feature
DefaultAssay(seurat_object_CD4) <- "ADT"
feature_plots_list_ADT <- lapply(adt_specific_markers, create_feature_plot)

DefaultAssay(seurat_object_CD4) <- "RNA"
feature_plots_list_RNA <- lapply(rna_specific_markers, create_feature_plot)

feature_plots_all <- c(feature_plots_list_ADT, feature_plots_list_RNA)
```

```{r}
feature_grid_plot <- plot_grid(plotlist = feature_plots_all, ncol = 4)
```



# CD8

# 1.A/B Splitting into CD8 Seurat Object

```
DefaultAssay(seurat_object) <- "ADT"
Idents(seurat_object) <- "RNA_clusters"

seurat_object_CD8 <- subset(x = seurat_object, subset = CD4 < 0.75)
#seurat_object_CD8 <- subset(x = seurat_object, subset = CD8A > 0.75 & CD4 < 0.75)
#saveRDS(seurat_object_CD4, file = paste0(data_dir, "/CD4_data/seurat_object_CD4.rds"))
```

```{r}
seurat_object_CD8 <- readRDS(file = paste0(data_dir, "/CD8_data/seurat_object_CD8.rds"))
```


# 1.C Dimension Reduction and Normalization

```
DefaultAssay(seurat_object_CD8) <- "RNA"

seurat_object_CD8 <- seurat_object_CD8 %>% NormalizeData(verbose = FALSE) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% ScaleData(verbose = TRUE)
seurat_object_CD8 <- seurat_object_CD8 %>% RunPCA(pc.genes = seurat_object_CD8@var.genes, npcs = 50, verbose = FALSE) %>%  RunUMAP(dims = 1:15, reduction.name = "umap_cd8_") %>% RunTSNE(dims = 1:15, reduction.name = "tsne_cd8_") %>% FindNeighbors(dims = 1:15,compute.SNN=TRUE) %>% FindClusters(resolution = c(0.1,0.2,0.4,0.6,0.8,1), cluster.name = "cd8_") %>% DoopenTSNE(reduction_save = "openTSNE_cd8_",dims_use = 1:15) %>% DoForceAtlas2(reduction_use = "pca", reduction_save = "fa2_cd8_", dims_use = 1:15)
```

```{r}
ElbowPlot(seurat_object_CD8, ndims = 50, reduction = "pca") + ggtitle("CD8 PCA Elbow Plot")
```

# 1.D Visualizing New CD8 clusters in new Dimension Reduction

# Split by Infection
```{r}
CD8_colors <- c("#865640",paletteer_d("ggthemes::calc"))
```

```{r}
Idents(object = seurat_object_CD8) <- "RNA_snn_res.0.6"
DimPlot(seurat_object_CD8, reduction = "pca", label = FALSE, repel = TRUE, split.by = "Infection", cols = CD8_colors)
DimPlot(seurat_object_CD8, reduction = "umap_cd8_", label = FALSE, repel = TRUE, split.by = "Infection", cols = CD8_colors)
DimPlot(seurat_object_CD8, reduction = "tsne_cd8_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD8, reduction = "openTSNE_cd8_", label = FALSE, repel = TRUE, split.by = "Infection")
DimPlot(seurat_object_CD8, reduction = "fa2_cd8_", label = FALSE, repel = TRUE, split.by = "Infection")
```

```
seurat_object_CD8$CD8_clust <- seurat_object_CD8$RNA_snn_res.0.6
```

# 1.G CD8 Cluster-Organ Frequency FIGURE 3 P2

```{r}
df_cd8_mouse <- table(seurat_object_CD8$Mouse_New, seurat_object_CD8$CD8_clust_star, seurat_object_CD8$Organ)
df_cd8_organ <- table(seurat_object_CD8$CD8_clust_star, seurat_object_CD8$Organ)

temp <- as.data.frame(melt(df_cd8_organ))
colnames(temp) <- c("Cluster","Organ", "Freq")

temp <- temp %>%
  group_by(Cluster) %>%
  mutate(Organ_Cluster_perc = Freq / sum(Freq) * 100)

#Fix the factor for visualization
temp$Cluster <- factor(temp$Cluster, levels =  c("1*", "8*", "9*", "7*", "4*", "6*", "3*", "0*", "5*", "2*"))

temp$Total_cells <- sum(temp$Freq)

temp <- temp %>%
  group_by(Cluster) %>% mutate(Cluster_perc = sum(Freq) / Total_cells * 100)

```

```{r}
organ_perc1_CD8 <- ggplot(temp, aes(x = Cluster, y = Organ_Cluster_perc, fill = Organ)) +
    geom_bar(stat = "identity", position = "fill") + 
    labs(title = "CD8 Clusters - Organ Proportion",
         x = "CD8 Cluster",
         y = "Proportion of each Cluster",
         fill = "Organ") + #theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x=element_blank(), legend.position = "none") + geom_label_repel(aes(label = sprintf("%.1f%%", Organ_Cluster_perc)), direction = ("y"),
               position = position_fill(vjust = 0.5), size = 3)
```

```{r}
cluster_sum <- aggregate(Cluster_perc ~ Cluster, data = temp, sum)

organ_perc2_CD8 <- ggplot(temp, aes(x = Cluster, y = Freq, fill = Organ)) +
    geom_bar(stat = "identity") + 
    #geom_text(data = Cluster_perc, aes(x = Cluster, y = max(temp$Freq) + 50, label = paste0(round(Cluster_perc, 1), "%")),
         #     position = position_dodge(width = 0.9), color = "black", size = 3) +
    labs(title = "CD8 Clusters - Total Number by Organ",
         x = "CD8 Cluster",
         y = "Number of Cells",
         fill = "Organ") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
```

```{r}
blank <- ggplot() + theme_void()

organ_perc3_CD8 <- DimPlot(seurat_object_CD8, group.by = "Organ", reduction = "tsne_cd8_") + ggtitle("CD8 TSNE by Organ") + NoLegend()

cd8_organ <- cowplot::plot_grid(organ_perc1_CD8,organ_perc2_CD8,organ_perc3_CD8,  ncol = 1, labels = c("D", "E", "F"))
```

# 1.H Double Negative Cells in CD8 Object

```{r}
DefaultAssay(seurat_object) <- "ADT"

# Double Negative Seurat
seurat_object_double_neg <- subset(x = seurat_object, subset = CD8A < 0.75 & CD4 < 0.75)
double_neg_barcodes <- seurat_object_double_neg$cell_barcode

# Add Metadata
seurat_object_CD8$double_neg <- seurat_object_CD8$cell_barcode %in% double_neg_barcodes
table(seurat_object_CD8$double_neg)

levels <- c("CD8+", "Double Negative")
my_factor <- as.factor(seurat_object_CD8$double_neg)
levels(my_factor) <- levels
table(my_factor)

seurat_object_CD8$double_neg <- my_factor

dubneg_plot <- DimPlot(seurat_object_CD8, reduction = "tsne_cd8_", group.by = "CD8_clust_star", cols = CD8_colors, split.by = "double_neg") + ggtitle("CD4-/CD8- Cells") + NoLegend() +xlim(-40,45) + ylim(-40,45)
dubneg_plot <- LabelClusters(plot = dubneg_plot, id = "CD8_clust_star", clusters = c("0*", "1*", "2*", "3*", "4*", "5*", "6*", "7*", "8*", "9*"), color = c("white", "white", "black", "black", "black", "white", "black", "white", "black", "white", "white", "white", "black", "black", "black", "white", "black", "white", "black", "white"), repel = TRUE, box = TRUE)

blank <- ggplot() + theme_void()

dubneg_plot2 <- cowplot::plot_grid(dubneg_plot, blank, rel_heights = c(1, 0.1), ncol = 1)
```

# Double Negative by Organ
```{r}
organ_table <- table(seurat_object_CD8$double_neg, seurat_object_CD8$Organ)
barplot(organ_table, main = "CD4-/CD8- Cell Organ Distribution", xlab = "Organ", ylab = "Frequency", col = "skyblue")

organ_infection_table <- table(seurat_object_CD8$Organ, seurat_object_CD8$Infection, seurat_object_CD8$double_neg)
organ_infection_df <- as.data.frame(organ_infection_table)
colnames(organ_infection_df) <- c("Organ", "Infection", "double_neg", "double_neg_count")

# Calculate percentage
organ_infection_df <- organ_infection_df %>%
  group_by(Organ, Infection) %>%
  mutate(percentage = double_neg_count / sum(double_neg_count) * 100)

# Create barplot
organ_dub_neg_plot <- ggplot(organ_infection_df, aes(x = Infection, y = percentage, fill = as.factor(double_neg))) +
  geom_bar(stat = "identity") +
  geom_label(aes(label = sprintf("%.1f%%", percentage), color = as.factor(double_neg)),
               position = position_stack(vjust = 0.5), size = 3) +
  geom_text(aes(label = sprintf("%.1f%%", percentage),
                color = as.factor(double_neg)),
            position = position_stack(vjust = 0.5),
            size = 3, show.legend = FALSE) +
  labs(title = "Percentage Double Negative cells (CD4-/CD8-) by Organ and Infection",
       x = "Infection",
       y = "Percentage",
       fill = "Surface Marker Expression") +
  scale_fill_manual(values = c("Double Negative" = "red", "CD8+" = "blue")) +
  scale_color_manual(values = c("Double Negative" = "black", "CD8+" = "white")) +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_grid(Organ ~ ., scales = "free_y") + guides(color="none")

```

# Double Negative by Clust

```{r}
double_neg_perc <- as.data.frame(table(seurat_object_CD8$double_neg, seurat_object_CD8$CD8_clust_star))
colnames(double_neg_perc) <- c("DoubleNeg", "Cluster", "Freq")

double_neg_perc <- double_neg_perc %>%
  group_by(Cluster) %>%
  mutate(percentage = Freq / sum(Freq) * 100)

double_neg_perc$Cluster <- factor(double_neg_perc$Cluster, levels = c("5*", "3*", "2*", "6*","8*", "4*", "9*", "7*", "0*", "1*"))

# Create ggplot bar graph
dub_neg_clust <- ggplot(double_neg_perc, aes(x = as.factor(Cluster), y = percentage, fill = as.factor(DoubleNeg))) +
    geom_bar(stat = "identity") +
    geom_label(aes(label = sprintf("%.1f%%", percentage), color = as.factor(DoubleNeg)),
               position = position_stack(vjust = 0.5), size = 3) +
    geom_text(aes(label = sprintf("%.1f%%", percentage), color = as.factor(DoubleNeg)),
              position = position_stack(vjust = 0.5), size = 3, show.legend = FALSE) +  
    labs(title = "Percentage of Double Negative (CD4-/CD8-) cells in each cluster",
         x = "CD8 Cluster",
         y = "Percentage", fill = "Surface Marker Expr") +
    scale_fill_manual(values = c("Double Negative" = "red", "CD8+" = "blue")) +
    scale_color_manual(values = c("Double Negative" = "black", "CD8+" = "white")) +
    theme_minimal() +
    theme(legend.position = "bottom") + guides(fill = guide_legend(override.aes = list(color = c("white", "black")) ))

dub_neg_clust <- ggplot(double_neg_perc, aes(x = as.factor(Cluster), y = percentage, fill = as.factor(DoubleNeg))) +
    geom_bar(stat = "identity") +
    geom_label(aes(label = sprintf("%.1f%%", percentage), color = as.factor(DoubleNeg)),
               position = position_stack(vjust = 0.5), size = 3) +
    geom_text(aes(label = sprintf("%.1f%%", percentage), color = as.factor(DoubleNeg)),
              position = position_stack(vjust = 0.5), size = 3, show.legend = FALSE) +  
    labs(title = "Percentage of Double Negative (CD4-/CD8-) cells in each cluster",
         x = "CD8 Cluster",
         y = "Percentage", fill = "Surface Marker Expr") +
    scale_fill_manual(values = c("Double Negative" = "red", "CD8+" = "blue"), guide = "legend") +
    scale_color_manual(values = c("Double Negative" = "black", "CD8+" = "white"), guide = "legend") +
    theme_minimal() +
    theme(legend.position = "bottom")

```


# Figure 1 - CD4 Split and New Clust

```{r}
DefaultAssay(seurat_object) <- "ADT"
#gfeat <- FeaturePlot(seurat_object, reduction = "umap_both_", features = c("CD4", "CD8A", "CD62L", "CD44"), combine = FALSE)
DefaultAssay(seurat_object) <- "ADT"
gfeat <- FeaturePlot(seurat_object, reduction = "tsne_both_", features = c("CD4", "CD8A", "CD44", "CD62L"), combine = FALSE, ncol = 2)

for(i in 1:length(gfeat)) {
  gfeat[[i]] <- gfeat[[i]] + NoLegend()
}

fig1a <- cowplot::plot_grid(plotlist = gfeat)

fig1b <- FeatureScatter(seurat_object, feature1 = "CD4", feature2 = "CD8A") + geom_vline(xintercept = 0.75, linetype = "dashed", color = "red", size = 1.5) + NoLegend()

# Cluster numbers taken from this file post- processing CD4 and CD8
all_clust_num <- readRDS(file = paste0(data_dir, "/all_clust_numbers.rds"))
merged_data <- merge(seurat_object@meta.data, all_clust_num, by.x = "cell_barcode", by.y = "barcodes", all.x = TRUE)

seurat_object$clust_num <- merged_data$clust_num

seurat_object$clust_num <- factor(seurat_object$clust_num, levels = c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "0*", "1*", "2*", "3*", "4*", "5*", "6*", "7*", "8*", "9*"))

vln4 <- VlnPlot(seurat_object, features = c("CD4"), group.by = "clust_num",pt.size = 0) +  geom_hline(yintercept = 0.75, linetype = "dashed", color = "red", size = 1.5) + NoLegend() +   theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank())
vln8 <- VlnPlot(seurat_object, features = c("CD8A"), group.by = "clust_num",pt.size = 0) + NoLegend() +   theme(axis.title.y=element_blank())

fig1cd <- cowplot::plot_grid(vln4, vln8, labels = c('C','D'), ncol = 1, rel_heights = c(0.9,1))

fig1bcd <- cowplot::plot_grid(fig1b, fig1cd , labels = c('B',''), ncol = 1, rel_heights = c(0.5,1))

fig1abcd <- cowplot::plot_grid(fig1a, fig1bcd, ncol =2, labels = c("A", ""))

#VlnPlot(seurat_object, features = "CD4") + geom_vline(xintercept = 0.75, linetype = "dashed", color = "red", size = 1.5)
```


```{r}

cd4_plot <- DimPlot(seurat_object_CD4, reduction = "tsne_cd4_", group.by = "CD4_clust", cols = CD4_colors) + ggtitle("CD4+ Cells") + NoLegend()
cd4_plot <- LabelClusters(plot = cd4_plot, id = "CD4_clust", clusters = c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16"), color = c("white", "black", "black", "black", "black", "white", "black", "black", "black", "white", "black", "white", "black", "white", "black", "white", "black"), repel = TRUE, box = TRUE)

cd8_plot <- DimPlot(seurat_object_CD8, reduction = "tsne_cd8_", group.by = "CD8_clust_star", cols = CD8_colors) + ggtitle("CD8+ and CD4- Cells") + NoLegend()
cd8_plot <- LabelClusters(plot = cd8_plot, id = "CD8_clust_star", clusters = c("0*", "1*", "2*", "3*", "4*", "5*", "6*", "7*", "8*", "9*"), color = c("white", "white", "black", "black", "black", "white", "black", "white", "black", "white"), repel = TRUE, box = TRUE)

fig1ef <- cowplot::plot_grid(cd4_plot, cd8_plot, nrow = 1, labels = c("E", "F"))
```

```{r}
fig1 <- cowplot::plot_grid(fig1abcd, fig1ef, ncol = 1, rel_heights = c(1,0.6))
```

```{r}
fig_directory <- "/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Candida_Migratory_TCR_Project/Candida_Migratory_TCR/Figures_Tables/"

ggsave(filename = paste0(fig_directory, "figure_1_2.pdf"), plot = fig1, device = "pdf", height = 10, width = 8)

ggsave(filename = paste0(fig_directory, "figure_1_2.jpeg"), plot = fig1, device = "jpeg", height = 10, width = 8)
```


# Figure 2 -  Double Negative


```{r}
dub_neg_perc <- cowplot::plot_grid(organ_dub_neg_plot , dub_neg_clust, rel_heights = c(0.85,1), labels = c("B", "C"), ncol =1)

fig2 <- cowplot::plot_grid(dubneg_plot2, dub_neg_perc, labels = c("A", ""))

ggsave(filename = paste0(fig_directory, "figure_2.pdf"), plot = fig2, device = "pdf", height = 10, width = 16)

ggsave(filename = paste0(fig_directory, "figure_2.jpeg"), plot = fig2, device = "jpeg", height = 10, width = 16)

```

# Figure 3 - Organ Distribution

```{r}
fig3 <- cowplot::plot_grid(cd4_organ, cd8_organ, ncol = 2)

ggsave(filename = paste0(fig_directory, "figure_3.pdf"), plot = fig3, device = "pdf", height = 14, width = 14)

ggsave(filename = paste0(fig_directory, "figure_3.jpeg"), plot = fig3, device = "jpeg", height = 14, width = 14)
```

# Supplemental Figure 3

```{r}
g4 <- ElbowPlot(seurat_object_CD4, ndims = 50) + geom_vline(xintercept = 15, linetype = "dashed", color = "red", size = 1.5) + ggtitle("CD4 Elbow Plot")
g8 <- ElbowPlot(seurat_object_CD8, ndims = 50) + geom_vline(xintercept = 15, linetype = "dashed", color = "red", size = 1.5) + ggtitle("CD8 Elbow Plot")

elbows <- cowplot::plot_grid(g4, g8, labels = c('A','B'))

pca1 <- DimPlot(seurat_object_CD4, reduction = "pca", group.by = "sample_name") + ggtitle("CD4 PCA") + NoLegend()
pca2 <- DimPlot(seurat_object_CD8, reduction = "pca", group.by = "sample_name") + ggtitle("CD8 PCA") + NoLegend()

pcas <- cowplot::plot_grid(pca1, pca2, labels = c('C','D'), rel_widths = c(1, 1), nrow =1) 

supp3 <- cowplot::plot_grid(elbows,pcas, ncol = 1)
```

```{r}
fig_directory <- "/Users/johnmorris/Desktop/Comp_Bio_MS_Weill_Cornell/Summer_2023/Betel Lab/Betel:Iliev Candida Project/Candida_Migratory_TCR_Project/Candida_Migratory_TCR/Figures_Tables/"

ggsave(filename = paste0(fig_directory, "supplemental_3.pdf"), plot = supp3, device = "pdf", height = 8, width = 10)

ggsave(filename = paste0(fig_directory, "supplemental_3.jpeg"), plot = supp3, device = "jpeg", height = 8, width = 10)
```

